Are rankings of financial analysts useful to investors?

### 2015-03-11 doc/paper/paper-knitr-APS.Rnw
edited paper to master

<<include=FALSE>>=
library(knitr)
opts_chunk$set(concordance=TRUE)
@

@@

\begin{abstract}
Some institutions issue rankings of financial analysts. Given that these rankings are \textit{ex-post} they may not be able useful to investors. In this paper  we show that trading strategies based on past rankings outperform a passive strategy.  In addition, we report that analysts issuing accurate price targets are more favorable to the investors  to follow rather than analysts issuing accurate EPS forecasts. In addition, we report that the performance of trade strategies based upon price target accuracy is superior one  based upon EPS forecast accuracy
%Finally, we show that a strategy based on the perfect foresight of analysts' performance gains the highest cumulative return.
\end{abstract}

@@

In recent years, some institutions have been very active in publishing and  selling  rankings of financial analysts. Some rankings  are based on privately held surveys of  buy-side analysts (e.g., the Institutional Investor's rankings of the All-America Research Teams\footnote{\url{http://www.institutionalinvestor.com/Research/4560/First-Team.html}} and Bloomberg's America's Best Stock Analysts\footnote{\sloppy \url{http://www.bloomberg.com/news/2013-08-14/jpmorgan-top-stock-picker-with-equities-out-of-lockstep.html}}); others are based on the performance of sell-side analysts (ThomsonReuters' top StarMine analysts\footnote{\url{http://excellence.thomsonreuters.com/award/starmine}}). In any of these cases, the rankings aim at identifying the top analysts. However, aside from personal acknowledgment among peers, it is still arguable whether these are useful \citep{desai2000ass} or are merely ``popularity contests" \citep{emery2009}. %In this paper, we support the argument that the top-ranked analysts, indeed, have stock picking skills.

We show that the top ranked analysts have stock picking skills. The contributions of our research is fourfold. First, we develop a trading strategy that transforms the rankings of financial analysts into inputs for the Black-Litterman model \citep{black1992}. Second, we show that annualized cumulative returns generated by some trading strategies  based upon analysts' rankings outperform the passive strategy (e.g., buy-and-hold the general stock market index). Third, we show that the strategy based upon the perfect foresight of rankings yields the highest cumulative annualized return. Fourth, we find that investors  are better off following analysts that issue the most accurate target prices, rather than those that issue the most accurate EPS forecasts.

@@

Yet, several authors have since stressed that  there are information-gathering costs and information is not immediately reflected on prices  \citep{grossman1980iie}. As such, prices may not  reflect all the available information at all time because if this were the case, those who spent resources to collect and analyze   information would not have an incentive to do it because there would not get any compensation for it.

@@

Some authors show that that financial analysts' recommendations create value to investors \citep{womack1996,barber2001}\footnote{\cite{womack1996} finds that  post-recommendation excess returns are not mean-reverting, but are significant and in the direction forecast by the analysts. \cite{barber2001} finds that over the period of 1986-1996 a portfolio of stocks with the most (least) favorable consensus analyst recommendations yields an average abnormal return of 4.13 (-4.91)\%.}. Assuming that some analysts produce valuable advice it make sense to rank analysts based on the accuracy of their recommendations.

StarMine rankings are based on financial analysts accuracy either on price target or EPS forecasts. To rank analysts based on EPS forecasts, StarMine developed a proprietary metric called a Single-stock Estimating Score (SES). This score measures ``... [a] relative accuracy; that is, analysts are compared against their peers. An analyst's SES can range from 0 to 100, with 50 representing the average analyst. To get a score higher than 50, an analyst must make estimates that are both significantly different from and more accurate than other analysts' estimates"\footnote{\url{http://excellence.thomsonreuters.com/award/starmine?award=Analyst+Awards&award_group=Overall+Analyst+Awards}}.

@@

The goal of our study is to evaluate if and how rankings  add value to investors.  With this purpose, we develop several sets of active trading strategies, selecting the stocks most favored by analysts. The first strategy is based on the consensus estimate (giving equal weights to analysts' recommendations). The second set of strategies takes in consideration the analysts' target price and EPS accuracy ranks to form ``smart estimates". For the latter set of strategies, we analyse different time information sets of the accuracy of the analysts.

We compare the performance of the strategies based upon two types of rankings (target price and EPS forecast accuracy). By doing this, we address the ongoing debate in the literature on whether analysts, when issue the target price reports, rely on simple growth-based models or use more complex models (such as the residual income model of \cite{ohlson1995}). For example, \cite{bradshaw2004} suggests that analysts' EPS forecasts are consistent with their price targets and that analysts use growth models based on EPS forecasts to estimate stocks target prices. Differently, \cite{simon2011} argue that the most accurate analysts rely on more complex models in setting their price targets.

A further major development in the theoretical accounting literature on equity valuation models is the abnormal earnings growth (AEG) model of \cite{ohlson2005}, which relates share price to the level of expected earnings per share\footnote{In words, abnormal growth in EPS is assumed to be the difference between expected EPS for the period less last period’s EPS invested at the cost of equity.}. In short, this model justifies a price earnings ratio for companies where next period’s earnings are expected to grow over and above the normal rate.

@@

\section{Trading Strategies}
%- BL model
%- Information sets
\label{sec:trading}
Our trading strategy use the framework for  active portfolio management proposed by \cite{black1992}.  The model incorporates ``views" in a CAPM framework, forming optimal portfolios in a mean-variance optimization setting. ``Views" are expectations on individual stocks future performance by investors/analysts.

@@

To proxy expected returns we compare the analysts' 12-month target prices (TP) with today's stock price. Confidence $\Omega$ for stock is based on variation  across analysts which is similar to the measure of dispersion in analysts' opinions outlined in \cite{diether2002}.

@@

\item Using the market price information available at the last day of quarter $t-1$, we obtain the market implicit returns for each stock $s$,  and the variance/co-variance matrix across stocks;

@@

For the strategies that weight the analysts' estimates of expected return, we use analysts' rankings and the weight of each analyst $j$ is based on his/her rank such that the top analyst has the weight of 1 and then the weights diminish as the rank increases.

@@

As described above, we use target price and EPS rankings.

\subsubsection{Target Price ranking}

Analysts are ranked on the basis of Proportional Mean Absolute Forecast Error (PMAFE) that measures the accuracy of a forecast  \citep{clement1999,brown2001,ertimur2007}. First,  we define the forecast daily error  $FE_{i,j}$ as the absolute value of the difference between analysts' target price $TP_{j}$ and the daily stock price $P$ for each stock $s$:

\begin{equation}
\label{dfe}
FE_{i,s}^{TP}=|{P_{s}-TP_{j,s}}|
\end{equation}
The PMAFE is given as:
\begin{equation}
\label{tp:pmafe}
PMAFE_{j,s}^{TP}=\frac{FE_{i,s}^{TP}}{\overline{FE_{s}^{TP}}}
\end{equation}
where $\overline{{FE}_{t,s}^{TP}}$ is the average over quarter forecasting error. The target price is fixed over the quarter unless it gets revised.

The rank  that enters equation (\ref{eq:weight}) is average analyst's $PMAFE^{TP}$ per quarter:
\begin{equation}
\overline{PMAFE_{j,s}^{TP}}=\frac{1}{T} \sum_{t=1}^{T} PMAFE_{j,t,s}^{TP}
\end{equation}

\begin{equation}
\label{tp:rank}
rank_{j,s}=\rank_{j=1}^{N} \left\{ \overline{PMAFE_{j,s}^{TP}} \right\}
\end{equation}
where $T$ is the number of trading days in a quarter. Figure \ref{fig:example} shows an example.

\subsubsection{EPS ranking} 
To compute the EPS rankings, we apply the same procedure as above:
\begin{equation}
FE_{j,s}^{EPS}=|{ACT_{s}-PRED_{j,s}}|
\end{equation}
\begin{equation}
PMAFE_{j,s}^{EPS}= \frac{FE_{j,s}^{EPS}}{\overline{FE_{s}^{EPS}}}
\end{equation}
\begin{equation}
\label{eps:rank}
rank_{j,s}=\rank_{j=1}^{N} \left\{ PMAFE_{j,s}^{EPS} \right\} 
\end{equation}
where $ACT_{s}$ and $PRED_{j,s}$ are the actual and  analyst $j$ forecast EPS respectively for each stock $s$.

\subsection{Defining the confidence of expected returns $\Omega$}
\label{def-omega}
The confidence of $Q$ are given by the coefficient of variation of forecasting errors:

\begin{equation}
CV = \frac{\sigma (FE)}{\overline{FE}}
\end{equation}
where $\sigma$ and $\overline{FE}$ are the standard deviation and the mean of the analysts' forecasting error for either TP or EPS. A low value of $CV$ reflects consensual estimates of either future prices or EPS.

@@

To proceed with the trading strategy, we need to establish which information we going to use to build the rankings that will be used to compute the weighted return estimates. Different analysts' ranks are obtained  if we select different time horizons. If use only recent information, we focus on recent performance of the analysts. This, of course, is more sensitive to unique episodes (e.g., a quarter, which has been surprisingly good or bad). If, alternatively, we opt to incorporate all analyst performance, the ranking is less affected by such events, yet, it may not reflect the current abilities of the analyst. In this paper, we use two information sets: the first uses only the  information about the analyst' performance in period $t-1$; the second, uses all the available  information for that particular analyst. We call the former the \naive{} set and the latter the \default{} set.

In addition to these sets,  we also create a hypothetical scenario that assumes we know the future information about the analyst' accuracy performance which would only be available at the end of $t$. In this setting, we are assuming that we could  perfectly predict the true analysts' ranking in advance. This represents the perfect foresight strategy; therefore, it serves as a reference point to the other trading strategies. We call this approach as the \tr{} set.

We formalize these sets as follows. 
\begin{itemize}
\item  the \tr{} %-- a perfect foresight information:
\begin{equation}
\label{q:true}
\widehat{Q_{.,t}}=Q_{.,t}
\end{equation}

\item  the \naive{} % -- $t-$ information:
\begin{equation}
\label{q:naive}
\widehat{Q_{.,t}}=Q_{.,t-1}
\end{equation}

\item  the \default{} %-- the entire history of analysts performance:
%\begin{itemize}
%\item ranking based views:
\begin{equation}
\label{q:default}
\widehat{Q_{.,t}} = \frac{1}{T} \sum_{t=1}^{T} Q_{.,t}
\end{equation}
%\item consensus based views:
%\begin{equation}
%\label{default.cons}
%\widehat{consQ_{t,s}} = \frac{1}{T} \sum_{T=1}^{T-1} consQ_{t,s}
%\end{equation}
%\end{itemize}
\end{itemize}
where the $.$ subscript corresponds to either $Q_{cons,s}$ (eq. \ref{consq}) or $Q_{rank,s}$ (eq. \ref{rankq}).

@@

The total number of brokers\footnote{We use words ``analyst" and ``broker" interchangeably.} in TP dataset is \Sexpr{q.data[!is.na(b.view),.N,by=Broker][,.N]}, covering \Sexpr{q.data[!is.na(b.view),.N,by=Stock][,.N]} S\&P500 stocks. Given the fact that financial analysts commonly issue TP on annual basis\footnote{According to Wharton Research Data Services (WRDS), 92.33\% of all price targets reported in I/B/E/S have a 12-month horizon \citep{glushkov2009}.}, we assume that analysts keep their TP forecasts valid for one calendar year unless it is revised. After one year we assume that TP recommendation expires.

Consistent with other studies on analysts' expected returns that work with price targets  \citep{bradshaw2002,brav2003,da2011}, we truncate the sample of $TP/P-1$ at the \Sexpr{ trunk.percent *100}\textsuperscript{th} percentile (values below \Sexpr{round(q.data[,min(trunk.view,na.rm=T)],3)}) and at the \Sexpr{(1-trunk.percent)*100}\textsuperscript{th} percentile (values above \Sexpr{ round(q.data[,max(trunk.view,na.rm=T)],3)}). This is done due to occurrence of the extreme values. Most of these extreme values are driven by misalignment errors found on I/B/E/S data\footnote{We found some differences between the  DataStream and I/B/E/S the databases. In some cases the stock-splits and the dividends were not properly adjusted.}. After trimming, the number of observations ($\mathrm{Stock} \times \mathrm{Broker} \times  \mathrm{Quarter}$) is reduced  from \Sexpr{ prettyNum(q.data[,.N], big.mark=' ')} to \Sexpr{ prettyNum(q.data[!is.na(trunk.view),.N],big.mark=' ')}.

@@

Table (\ref{tab:ret-stat}) shows the distribution of the final sample of target price and EPS data. Panel A shows the number of quarterly target prices per stock. For the sample period (1999-2009), we report \Sexpr{prettyNum(core.dt[,.N,by=list(Stock,q.id)][,.N],big.mark=' ')} stock-quarter observations. Each stock had on average of \Sexpr{round(core.dt[,.N,by=list(Stock,q.id)][,mean(N)],2)} quarterly target price reports. Panel B  presents similar statistics of the number of EPS forecasts. In total, the EPS forecast sample has \Sexpr{prettyNum(complete.dt[,.N,by=list(Stock,q.id)][,.N],big.mark=' ')} stock-quarter observations. The average number of quarterly forecasts is \Sexpr{round(complete.dt[,.N,by=list(Stock,q.id)][,mean(N)],2)}.

We apply the ranking procedure outline in section \ref{def-q} on two datasets. For target price rankings, we use the average per quarter daily errors within one quarter as the measure of analysts' forecasting ability (eq. \ref{tp:pmafe}).

Table \ref{tab:example} and figure \ref{fig:example} illustrate an example of how we estimate the \textit{PMAFE}. Four analysts had valid target price for Amazon for second quarter of 1999. We plot the daily Amazon price against the brokers' target prices. Table \ref{tab:example} shows the resulting TP and EPS ranking. On the bases of the average daily errors, LEGG is the most accurate in forecasting stock price and  DLJ is the least accurate. For the EPS case (panel B), PACCREST is the most accurate in EPS forecasting and RBRTSON is the least.

@@

We convert the rankings into scores as follows:
\begin{equation}
\label{eq:score}
score_{j,s}=\frac{rank_{j,s}}{\max{rank}}
\end{equation}

To get the cross-sectional values of scores across different stocks, we take the average of $score_{j,s}$
\begin{equation}
\label{eq:mean-score}
score_{j}= \frac{1}{k} \sum_{s=1}^{k} score_{j,s}
\end{equation}
where $k$ is number of stocks for which an analyst $j$ had rank.

Table (\ref{tab:rank-stat}) shows a contingency analysis of the ranks. 
%Chi square statistics include/ expected frequiency test
%Journ. of Fin economics 
Panel A shows the dynamics of each tercile for rankings based on the accuracy in target prices. We observe that analysts exhibit strong ranking consistency as, on average, they stay at the same tercile after one quarter and even after one year. The case of EPS ranking shows similar results (panel B). This analysis confirms recent finding of \cite{hilary2013} on analyst forecast consistency.

@@

\subsection{$Q$ descriptive statistics}

Tables \ref{tab:view-stat} -- \ref{tab:view-stat-default} present the descriptive statistics of analysts' expected returns conditional on the information sets. The expected returns  are computed comparing TP estimates with actual price. To inform the strategies we obtain rank-weighted estimates where weights are given either by the TP or the EPS ranks.

\cite{bradshaw2002} reports that analysts' average expected returns for the period of 2000--2009 of 206 brokers 24\%, in \cite{da2011} these average expected return post 40\% for the period of 1996--2004, \cite{zhou2013} find on average  96\% for the sample period of 2000--2009 reports. The figures suggest that analyst overly optimistic.
%these values from the historical perspective, i.e., how the values of expected stock returns go inline with historical stock returns. \cite{bodie2009} show that arithimatic average rate of return for the U.S. large stocks (S\&P 500) for the period of 1926--2005 is 10.17\% and the average rate of excess return is 8.39\% with the risk premium estimated 6--8\%. While it is not the best idea to extrapolate the historical values, still, we can say that the expected stocks return should be around 14--16\%. Clearly, the values presented in the literature shows that analysts are very optimistic in issuing target price reports.

Panel A show the statistics for the consensus expectations as defined in eq. (\ref{consq}). The average consensus among the analysts about stock expected return for \tr{}, \naive{} and \default{} information sets are \Sexpr{round(stat.nr[year=='Total'&Method=='true',mean],3)*100 }\%, \Sexpr{round(stat.nr[year=='Total'&Method=='naive',mean],3)*100 }\%, and \Sexpr{round(stat.nr[year=='Total'&Method=='default',mean],3)*100 }\% respectively. Panel B of the table shows the TP accuracy weighted average expected returns (\Sexpr{ round(stat.pt.view[year=='Total'&Method=='true',mean],3)*100}\%, \Sexpr{ round(stat.pt.view[year=='Total'&Method=='naive',mean],3)*100}\%, and \Sexpr{ round(stat.pt.view[year=='Total'&Method=='default',mean],3)*100}\%). Panel C shows the EPS based weighted expected returns averages (\Sexpr{ round(stat.eps.view[year=='Total'&Method=='true',mean],3)*100}\%, \Sexpr{ round(stat.eps.view[year=='Total'&Method=='naive',mean],3)*100}\%, and \Sexpr{ round(stat.eps.view[year=='Total'&Method=='default',mean],3)*100}\%).

Overall compared to the consensus the ranked weighted expected returns (smart estimates) are less optimistic. The \default{} information set shows higher values of expected returns considering that it includes the whole history of analysts' performance.

%obtained from the rankings based on the target price accuracy. We observe a reduction in average per stock expected return with the mean of \Sexpr{ round(rank.views.full.period['true','mean'],3)*100}\%. The case of rankings based on EPS forecasts (panel C) shows even more moderate average stock expected return of \Sexpr{ round(pt.eps.rank.views.full.period['true','mean'],3)*100}\%. Overall, compared to the consensus, both ranking-based methods exhibit less optimistic expected returns and are pretty close to the historical averages.

@@

We report results from trading strategies in table (\ref{tab:strategy}). We split the table into four panels: one for the passive strategy and three for each of the information availability sets.

>>>>

<<include=FALSE>>=
library(knitr)
opts_chunk$set(concordance=TRUE)
options(digits = 3)
@

@@

\begin{abstract}
Several institutions issue rankings based on the accuracy of price and EPS forecasts. Given that these rankings are \textit{ex-post} they may not be able useful to investors. In this paper  we show that trading strategies based on perfect foresight and past rankings outperform a passive strategy.  In addition, we report that investors are better off following analysts that issue  accurate price targets rather that those with accurate EPS forecasts. 
%Finally, we show that a strategy based on the perfect foresight of analysts' performance gains the highest cumulative return.
\end{abstract}

@@

In recent years, some institutions have been very active in publishing and  selling  rankings of financial analysts. Some rankings  are based on privately held surveys of  buy-side analysts (e.g., the Institutional Investor's rankings of the All-America Research Teams\footnote{\url{http://www.institutionalinvestor.com/Research/4560/First-Team.html}} and Bloomberg's America's Best Stock Analysts\footnote{\sloppy \url{http://www.bloomberg.com/news/2013-08-14/jpmorgan-top-stock-picker-with-equities-out-of-lockstep.html}}); others are based on the performance of sell-side analysts (ThomsonReuters' top StarMine analysts\footnote{\url{http://excellence.thomsonreuters.com/award/starmine}}). In any of these cases, the rankings aim at identifying the top analysts. However, aside from personal acknowledgment among peers, it is still arguable whether these are useful to investors \citep{desai2000ass} or are merely ``popularity contests" \citep{emery2009}. %In this paper, we support the argument that the top-ranked analysts, indeed, have stock picking skills.

We show that the top ranked analysts have stock picking skills. The contributions of our research is fourfold. First, we develop a trading strategy that transforms the rankings of financial analysts into inputs for the Black-Litterman model \citep{black1992}. Second, we show that annualized cumulative returns generated by some trading strategies  based upon analysts' rankings outperform a passive strategy (e.g., buy-and-hold the general stock market index). Third, we show that the strategy based upon the perfect foresight of rankings yields the highest cumulative annualized return. Fourth, we find that investors  are better off following analysts that issue the most accurate target prices, rather than those that issue the most accurate EPS forecasts.

@@

Yet, several authors have since stressed that  there are information-gathering costs and information is not immediately reflected on prices  \citep{grossman1980iie}. As such, prices may not  reflect all the available information at all time because if this were the case, those who spent resources to collect and analyze   information would not have an incentive to do it, because there would not get any compensation for it.

@@

Some authors show that financial analysts' recommendations create value to investors \citep{womack1996,barber2001}\footnote{\cite{womack1996} finds that  post-recommendation excess returns are not mean-reverting, but are significant and in the direction forecast by the analysts. \cite{barber2001} finds that over the period of 1986-1996 a portfolio of stocks with the most (least) favorable consensus analyst recommendations yields an average abnormal return of 4.13 (-4.91)\%.}. Assuming that some analysts produce valuable advice it makes sense to rank analysts based on the accuracy of their recommendations.

StarMine rankings are based on financial analysts' accuracy either on TP or EPS forecasts. To rank analysts based on EPS forecasts, StarMine developed a proprietary metric called a Single-stock Estimating Score (SES). This score measures ``... [a] relative accuracy; that is, analysts are compared against their peers. An analyst's SES can range from 0 to 100, with 50 representing the average analyst. To get a score higher than 50, an analyst must make estimates that are both significantly different from and more accurate than other analysts' estimates"\footnote{\url{http://excellence.thomsonreuters.com/award/starmine?award=Analyst+Awards&award_group=Overall+Analyst+Awards}}.

@@

The goal of our study is to evaluate if and how rankings  add value to investors.  With this purpose, we develop several sets of active trading strategies, selecting the stocks most favored by analysts. The first strategy is based on the consensus estimate (giving equal weights to analysts' recommendations). The second set of strategies takes in consideration the analysts' target price and EPS accuracy ranks to form ``smart estimates". For the latter set of strategies, we analyse different time information sets to define the accuracy of the analysts.

We compare the performance of the strategies based upon two types of rankings (target price and EPS forecast accuracy). By doing this, we indirectly address the ongoing debate in the literature on whether analysts, when issuing the target price reports, rely on simple growth-based models or use more complex models (such as the residual income model of \cite{ohlson1995}). For example, \cite{bradshaw2004} suggests that analysts' EPS forecasts are consistent with their price targets and that analysts use growth models based on EPS forecasts to estimate stocks target prices. Differently, \cite{simon2011} argue that the most accurate analysts rely on more complex models in setting their price targets. \footnote{A further major development in the theoretical accounting literature on equity valuation models is the abnormal earnings growth (AEG) model of \cite{ohlson2005}, which relates share price to the level of expected earnings per share.}

@@

\section{Trading Strategies}
%- BL model
%- Information sets
\label{sec:trading}
Our trading strategy uses the framework for  active portfolio management proposed by \cite{black1992}.  The model incorporates ``views" in a CAPM framework, forming optimal portfolios in a mean-variance optimization setting. ``Views" are expectations on individual stocks' future performance.

@@

To proxy expected returns we compare the analyst' 12-month target price (TP) with today's stock price. Confidence $\Omega$ for stock is based on variation  of forecasts across analysts which is similar to the measure of dispersion in analysts' opinions outlined in \cite{diether2002}.

@@

\item Using the market price information available at the last day of quarter $t-1$, we obtain the market implicit returns for each stock $s$,  and the variance/co-variance matrix;

@@

For the strategies that weight the analysts' estimates of expected return the weight of each analsyt $j$ is based on his/her rank such that the top analyst has the weight of 1 and then the weights diminish as the rank increases.

@@

As mentioned above, we use both target price and EPS accuracy rankings.

\subsubsection{Target Price ranking} 
%%% change s to i for stock
Analysts are ranked on the basis of Proportional Mean Absolute Forecast Error (PMAFE) that measures the accuracy of a forecast  \citep{clement1999,brown2001,ertimur2007}. First,  we define the forecast daily error  $FE_{j,s}$ as the absolute value of the difference between analyst' target price $TP_{j}$ and the daily stock price $P$ for each stock $s$:

\begin{equation}
\label{dfe}
FE_{j,s}^{TP}=|{P_{s}-TP_{j,s}}|
\end{equation}
The PMAFE is given as:
\begin{equation}
\label{tp:pmafe}
PMAFE_{j,s}^{TP}=\frac{FE_{j,s}^{TP}}{\overline{FE_{s}^{TP}}}
\end{equation}
where $\overline{{FE}_{s}^{TP}}$ is the average forecasting error accros analysts. The target price is fixed over the quarter unless it gets revised.

The rank  that enters equation (\ref{eq:weight}) is average analyst's $PMAFE^{TP}$ over a particular quarter:
\begin{equation}
\overline{PMAFE_{j,s}^{TP}}=\frac{1}{T} \sum_{t=1}^{T} PMAFE_{j,t,s}^{TP}
\end{equation}

\begin{equation}
\label{tp:rank}
rank_{j,s}=\rank_{j=1}^{N} \left\{ \overline{PMAFE_{j,s}^{TP}} \right\}
\end{equation}
where $T$ are the number of trading days in a quarter and $N$ is the number of brokers with a valid TP.  Figure \ref{fig:example} shows an example.

\subsubsection{EPS ranking} 
To compute the EPS rankings, we apply the same procedure as above:
\begin{equation}
FE_{j,s}^{EPS}=|{ACT_{s}-PRED_{j,s}}|
\end{equation}
\begin{equation}
PMAFE_{j,s}^{EPS}= \frac{FE_{j,s}^{EPS}}{\overline{FE_{s}^{EPS}}}
\end{equation}
\begin{equation}
\label{eps:rank}
rank_{j,s}=\rank_{j=1}^{N} \left\{ PMAFE_{j,s}^{EPS} \right\} 
\end{equation}
where $ACT_{s}$ and $PRED_{j,s}$ are the actual quarterly EPS and  analyst $j$'s EPS forecast for stock $s$.

\subsection{Defining the confidence of expected returns $\Omega$}
\label{def-omega}
The confidence of $Q$ is given by the coefficient of variation (CV) of forecasting errors. For each stock $s$ is given by:

\begin{equation}
CV_{s} = \frac{\sigma (FE)_s}{\overline{FE}_s}
\end{equation}
where $\sigma_s$ and $\overline{FE}_s$ are the standard deviation and the mean of the forecast errors across analsyts for either TP or EPS. A low value of $CV$ reflects consensual estimates of either future prices or EPS.

@@

To proceed with the trading strategy, we need to establish which information we  will be using to build the rankings. These rankgins will be the inputs to compute the weighted return estimates (``smart estimates"). Different analysts' ranks are obtained  if we select different time horizons. If we use only the most  recent information, we will capture the recent performance of the analysts. This, of course, is more sensitive to unique episodes (e.g., a quarter which has been surprisingly good or bad). If, alternatively, we opt to incorporate the entire analyst performance, the ranking is less affected by such events, yet it may not reflect the current analyst ability. We use two information sets: the first uses only the  information about the analyst' performance in period $t-1$; the second, uses all the available  information for that particular analyst. We call the former the \naive{} set and the latter the \default{} set.

In addition to these sets,  we also create a hypothetical scenario that assumes we antecipate perfectly the future analyst accuracy performance  that would only be available at the end of $t$.  This represents the perfect foresight strategy. Therefore, it serves a performance reference point to evaluate the other trading strategies. We call this the \tr{} set.

Formalizing information sets considerd are: 
\begin{itemize}
\item  the \tr{} set%-- a perfect foresight information:
\begin{equation}
\label{q:true}
\widehat{Q_{.,t}}=Q_{.,t}
\end{equation}

\item  the \naive{} set % -- $t-$ information:
\begin{equation}
\label{q:naive}
\widehat{Q_{.,t}}=Q_{.,t-1}
\end{equation}

\item  the \default{}  set%-- the entire history of analysts performance:
%\begin{itemize}
%\item ranking based views:
\begin{equation}
\label{q:default}
\widehat{Q_{.,t}} = \frac{1}{T} \sum_{t=1}^{T} Q_{.,t}
\end{equation}
%\item consensus based views:
%\begin{equation}
%\label{default.cons}
%\widehat{consQ_{t,s}} = \frac{1}{T} \sum_{T=1}^{T-1} consQ_{t,s}
%\end{equation}
%\end{itemize}
\end{itemize}
where the $.$ subscript corresponds to either $Q_{cons,s}$ (eq. \ref{consq}) or $Q_{rank,s}$ (eq. \ref{rankq}), i.e. this is consensus or smart estimate.

@@

Over the sample period, the total number of brokers\footnote{We use words ``analyst" and ``broker" interchangeably.} in TP dataset is \Sexpr{q.data[!is.na(b.view),.N,by=Broker][,.N]}, covering \Sexpr{q.data[!is.na(b.view),.N,by=Stock][,.N]} S\&P500 stocks. Given the fact that financial analysts commonly issue TP with the one year horizion \footnote{According to Wharton Research Data Services (WRDS), 92.33\% of all price targets reported in I/B/E/S have a 12-month horizon \citep{glushkov2009}.}, we assume that analysts keep their TP forecasts valid for one calendar year unless it is revised. After one year we assume that TP recommendation expires.

Consistent with other studies on analysts' expected returns that work with price targets  \citep{bradshaw2002,brav2003,da2011}, we truncate the sample of $TP/P-1$ at the \Sexpr{ trunk.percent *100}\textsuperscript{th} percentile (values below \Sexpr{q.data[,min(trunk.view,na.rm=T)]}) and at the \Sexpr{(1-trunk.percent)*100}\textsuperscript{th} percentile (values above \Sexpr{ round(q.data[,max(trunk.view,na.rm=T)],3)}). This is done due to occurrence of the extreme values. Most of these extreme values are driven by misalignment errors found on I/B/E/S data\footnote{We found some differences between the  DataStream and I/B/E/S the databases. In some cases the stock-splits and the dividends were not properly adjusted.}. After trimming, the number of observations ($\mathrm{Stock} \times \mathrm{Broker} \times  \mathrm{Quarter}$) is reduced  from \Sexpr{ prettyNum(q.data[,.N], big.mark=' ')} to \Sexpr{ prettyNum(q.data[!is.na(trunk.view),.N],big.mark=' ')}.

@@

Table (\ref{tab:ret-stat}) shows the distribution of the final sample of target price and EPS data. Panel A shows the number of quarterly target prices per stock. For the sample period (1999-2009), we report \Sexpr{prettyNum(core.dt[,.N,by=list(Stock,q.id)][,.N],big.mark=' ')} stock-quarter observations. Each stock had on average of \Sexpr{core.dt[,.N,by=list(Stock,q.id)][,mean(N)]} quarterly target price reports. Panel B  presents similar statistics of the number of EPS forecasts. In total, the EPS forecast sample has \Sexpr{prettyNum(complete.dt[,.N,by=list(Stock,q.id)][,.N],big.mark=' ')} stock-quarter observations. The average number of quarterly forecasts is \Sexpr{complete.dt[,.N,by=list(Stock,q.id)][,mean(N)]}.

We apply the ranking procedure outline in section \ref{def-q} to the two datasets. For target price rankings, we use the average daily errors within one quarter as the measure of analysts' forecasting ability (eq. \ref{tp:pmafe}).

Table \ref{tab:example} and figure \ref{fig:example} illustrate an example illustrate how we estimate the \textit{PMAFE}. Four analysts had valid target prices for Amazon for second quarter of 1999. We plot the daily Amazon price against the brokers' target prices. Table \ref{tab:example} shows the resulting TP and EPS rankings. On the bases of the average daily errors, LEGG is the most accurate in forecasting stock price and  DLJ is the least accurate. For the EPS case (panel B), PACCREST is the most accurate in EPS forecasting and RBRTSON is the least.

@@

Beforend, we convert the rankings into scores as follows:
\begin{equation}
\label{eq:score}
score_{j,s}=\frac{rank_{j,s}}{\max{rank_s}}
\end{equation}

To get the cross-sectional values of scores across different stocks, we take the average of $score_{j,s}$
\begin{equation}
\label{eq:mean-score}
\overline{score_{j}}= \frac{1}{k} \sum_{s=1}^{k} score_{j,s}
\end{equation}
where $k$ is number of stocks followed by a particular analyst $j$.

Table (\ref{tab:rank-stat}) shows a contingency analysis of the ranks. 
%Chi square statistics include/ expected frequiency test
%Journ. of Fin economics 
Panel A shows the dynamics of each tercile for rankings based on target price  accuracy. We observe that analysts exhibit strong ranking consistency as, on average, they stay at the same tercile after one quarter. Of the top (bottom) most accurate (inaccurate) analysts in the previous quarter 68\% (69\%) remain in that same tercile after one quarter. After one year the corresponding figures are lower respectively 47 and 49\% for the top and bottom terciles.

In the case of EPS ranking  (panel B) 52 and 57 (42 and 40) of the analsyts  remained in the top and bottom terciles after one quarter (year)  respectively.

These results are consistent with the recent findings of \cite{hilary2013} on analyst forecast consistency.

@@

\subsection{Views: descriptive statistics}

Tables \ref{tab:view-stat} to \ref{tab:view-stat-default} present the descriptive statistics of the analysts' expected returns conditional on the different information sets. Table 4 shows the \tr{}, table 5 shows the \naive{} and table 6 \defaul{}.

The expected returns  are computed comparing TP estimates with actual prices. To form the smart strategies we compute rank-weighted estimates where weights are given either by the TP or the EPS ranks.

\cite{bradshaw2002} reports analyst average expected returns for the period of 2000--2009 and 206 brokers of 24\%. \cite{da2011} report an average expected return of 40\% for the period of 1996--2004. \cite{zhou2013} finds an average expected return of 96\% for the sample period of 2000--2009. These figures suggest that analysts are overly optimistic.
%these values from the historical perspective, i.e., how the values of expected stock returns go inline with historical stock returns. \cite{bodie2009} show that arithimatic average rate of return for the U.S. large stocks (S\&P 500) for the period of 1926--2005 is 10.17\% and the average rate of excess return is 8.39\% with the risk premium estimated 6--8\%. While it is not the best idea to extrapolate the historical values, still, we can say that the expected stocks return should be around 14--16\%. Clearly, the values presented in the literature shows that analysts are very optimistic in issuing target price reports.

Panel A show the statistics for the consensus expectations as defined in eq. (\ref{consq}). The average consensus among the analysts on stock expected returns in the  \tr{}, \naive{} and \default{} information sets are, respectively, \Sexpr{stat.nr[year=='Total'&Method=='true',mean]*100 }\%, \Sexpr{stat.nr[year=='Total'&Method=='naive',mean]*100 }\%, and \Sexpr{stat.nr[year=='Total'&Method=='default',mean]*100 }\% respectively. Panel B of the table shows the TP accuracy weighted average expected returns (\Sexpr{stat.pt.view[year=='Total'&Method=='true',mean]*100}\%, \Sexpr{stat.pt.view[year=='Total'&Method=='naive',mean]*100}\%, and \Sexpr{stat.pt.view[year=='Total'&Method=='default',mean]*100}\%). Panel C shows the EPS based weighted expected returns averages (\Sexpr{stat.eps.view[year=='Total'&Method=='true',mean]*100}\%, \Sexpr{stat.eps.view[year=='Total'&Method=='naive',mean]*100}\%, and \Sexpr{stat.eps.view[year=='Total'&Method=='default',mean]*100}\%).

Overall compared to the consensus the ranked weighted expected returns (smart estimates) are less optimistic. The \default{} information set shows higher values of expected returns.
%obtained from the rankings based on the target price accuracy. We observe a reduction in average per stock expected return with the mean of \Sexpr{ round(rank.views.full.period['true','mean'],3)*100}\%. The case of rankings based on EPS forecasts (panel C) shows even more moderate average stock expected return of \Sexpr{ round(pt.eps.rank.views.full.period['true','mean'],3)*100}\%. Overall, compared to the consensus, both ranking-based methods exhibit less optimistic expected returns and are pretty close to the historical averages.

@@

We report the results from different trading strategies in table (\ref{tab:strategy}). We split the table into four panels. Panel A shows  the performance for  the passive (market) strategy \textit{Market}. Panels B to D compare the consensus , the TP rank weighted and the EPS rank weighted trading strategis for each of the information availability sets.

\subsection{Passive strategy}

The passive strategy generates an annualized cumulative return of \Sexpr{bl.results['Market',1,1]}\% with a Sharpe Ratio  of \Sexpr{bl.results['Market',1,'ann.sr']} over the period 1999-2001. The average number of stocks held per quarter was \Sexpr{bl.results['Market',1,'meanViews']} and the turnover ratio was \Sexpr{bl.results['Market',1,'Ave.TO']}, which reflects solely the inclusion and deletions  of the S\&P 500 constituent list.
